---
title: "dplyr_across_albert_rapp_example"
format: html
editor: visual
---

## 

|     |     |
|-----|-----|
|     |     |

Bad summarize

```{r}
library(tidyverse)
palmerpenguins::penguins |> 
  summarise(
    mean_bill_length = mean(bill_length_mm, na.rm = TRUE),
    mean_bill_depth = mean(bill_depth_mm, na.rm = TRUE),
    mean_flipper_length = mean(flipper_length_mm, na.rm = TRUE),
    mean_body_mass = mean(body_mass_g, na.rm = TRUE)
  )
```

Good summarize

```{r}
palmerpenguins::penguins |> 
  summarise(
    across(
      .cols = c(
        bill_length_mm, bill_depth_mm, 
        flipper_length_mm, body_mass_g
      ),
      .fns = \(x) mean(x, na.rm = TRUE)
    )
  )
```

Also, this works as per syntax documentation

```{r}
palmerpenguins::penguins |> 
  summarize(
    across(
      .cols = c(
        bill_length_mm, bill_depth_mm, 
        flipper_length_mm, body_mass_g
      ), 
      .fns = ~mean(.x, na.rm = TRUE)
    )
  )
```

Using multiple functions

```{r}
palmerpenguins::penguins |> 
  summarise(
    across(
      .cols = c(
        bill_length_mm, bill_depth_mm, 
        flipper_length_mm, body_mass_g
      ),
      .fns = c(
        \(x) mean(x, na.rm = TRUE),
        \(x) var(x, na.rm = TRUE)
      )
    )
  )
```

Also, we can use the followings

```{r}
palmerpenguins::penguins |> 
  summarise(
    across(
      .cols = c(
        bill_length_mm, bill_depth_mm, 
        flipper_length_mm, body_mass_g
      ),
      .fns = c(
        ~mean(.x, na.rm = TRUE), 
        ~var(.x, na.rm = TRUE)
      )
    )
  )
```

Better column names by naming function in vector

```{r}
palmerpenguins::penguins |> 
  summarise(
    across(
      .cols = c(
        bill_length_mm, bill_depth_mm, 
        flipper_length_mm, body_mass_g
      ),
      .fns = c(
        mean = \(x) mean(x, na.rm = TRUE),
        variance =\(x) var(x, na.rm = TRUE)
      )
    )
  )
```

OR

```{r}
palmerpenguins::penguins |> 
  summarise(
    across(
      .cols = c(
        bill_length_mm, bill_depth_mm, 
        flipper_length_mm, body_mass_g
      ),
      .fns = c(
       mean =  ~mean(.x, na.rm = TRUE), 
       variance =  ~var(.x, na.rm = TRUE)
      )
    )
  )
```

Better separator for pivot_longer()

```{r}
palmerpenguins::penguins |> 
  summarise(
    across(
      .cols = c(
        bill_length_mm, bill_depth_mm, 
        flipper_length_mm, body_mass_g
      ),
      .fns = c(
        mean = \(x) mean(x, na.rm = TRUE),
        variance =\(x) var(x, na.rm = TRUE)
      ),
      .names = '{.col}----{.fn}'
    )
  ) |> 
  pivot_longer(
    cols = everything(),
    names_sep = '----',
    names_to = c('variable', 'stat')
  )
```

Better approach for many columns using tidyselect helpers

```{r}
modeldata::ames |> 
  summarise(
    across(
      .cols = where(is.numeric),
      .fns = c(
        mean = \(x) mean(x, na.rm = TRUE),
        variance =\(x) var(x, na.rm = TRUE)
      ),
      .names = '{.col}----{.fn}'
    )
  ) |> 
  pivot_longer(
    cols = everything(),
    names_to = c('variable', 'stat'),
    names_sep = '----'
  )
```

Combine multiple select helpers

```{r}
modeldata::ames |> 
  summarise(
    across(
      .cols = where(is.numeric) & starts_with('Lot'),
      .fns = c(
        mean = \(x) mean(x, na.rm = TRUE),
        variance =\(x) var(x, na.rm = TRUE)
      ),
      .names = '{.col}----{.fn}'
    )
  ) |> 
  pivot_longer(
    cols = everything(),
    names_to = c('variable', 'stat'),
    names_sep = '----'
  )
```

You can use across() with mutate as well.

```{r}
modeldata::ames |> 
  mutate(
    across(
      .cols = matches('(Y|y)ear'),
      .fns = \(x) lubridate::make_date(year = x)
    )
  ) |> 
  select(where(lubridate::is.Date))
```

Bad summarise

```{r}
library(tidyverse)
palmerpenguins::penguins |> 
  summarise(
    mean_bill_length = mean(bill_length_mm, na.rm = TRUE),
    mean_bill_depth = mean(bill_depth_mm, na.rm = TRUE),
    mean_flipper_length = mean(flipper_length_mm, na.rm = TRUE),
    mean_body_mass = mean(body_mass_g, na.rm = TRUE)
  )
```

Good summarise

```{r}
palmerpenguins::penguins |> 
  summarise(
    across(
      .cols = c(
        bill_length_mm, bill_depth_mm, 
        flipper_length_mm, body_mass_g
      ),
      .fns = \(x) mean(x, na.rm = TRUE)
    )
  )
```

Using multiple functions

```{r}
palmerpenguins::penguins |> 
  summarise(
    across(
      .cols = c(
        bill_length_mm, bill_depth_mm, 
        flipper_length_mm, body_mass_g
      ),
      .fns = c(
        \(x) mean(x, na.rm = TRUE),
        \(x) var(x, na.rm = TRUE)
      )
    )
  )
```

Better column names by naming function in vector

```{r}
palmerpenguins::penguins |> 
  summarise(
    across(
      .cols = c(
        bill_length_mm, bill_depth_mm, 
        flipper_length_mm, body_mass_g
      ),
      .fns = c(
        mean = \(x) mean(x, na.rm = TRUE),
        variance =\(x) var(x, na.rm = TRUE)
      )
    )
  )
```

Better separator for pivot_longer()

```{r}
palmerpenguins::penguins |> 
  summarise(
    across(
      .cols = c(
        bill_length_mm, bill_depth_mm, 
        flipper_length_mm, body_mass_g
      ),
      .fns = c(
        mean = \(x) mean(x, na.rm = TRUE),
        variance =\(x) var(x, na.rm = TRUE)
      ),
      .names = '{.col}----{.fn}'
    )
  ) |> 
  pivot_longer(
    cols = everything(),
    names_sep = '----',
    names_to = c('variable', 'stat')
  )
```

Better approach for many columns using tidyselect helpers

```{r}
modeldata::ames |> 
  summarise(
    across(
      .cols = where(is.numeric),
      .fns = c(
        mean = \(x) mean(x, na.rm = TRUE),
        variance =\(x) var(x, na.rm = TRUE)
      ),
      .names = '{.col}----{.fn}'
    )
  ) |> 
  pivot_longer(
    cols = everything(),
    names_to = c('variable', 'stat'),
    names_sep = '----'
  )
```

Combine multiple select helpers

```{r}
modeldata::ames |> 
  summarise(
    across(
      .cols = where(is.numeric) & starts_with('Lot'),
      .fns = c(
        mean = \(x) mean(x, na.rm = TRUE),
        variance =\(x) var(x, na.rm = TRUE)
      ),
      .names = '{.col}----{.fn}'
    )
  ) |> 
  pivot_longer(
    cols = everything(),
    names_to = c('variable', 'stat'),
    names_sep = '----'
  )
```

You can use across() with mutate as well.

```{r}
modeldata::ames |> 
  mutate(
    across(
      .cols = matches('(Y|y)ear'),
      .fns = \(x) lubridate::make_date(year = x)
    )
  ) |> 
  select(where(lubridate::is.Date))
```

Bad summarise

```{r}
library(tidyverse)
palmerpenguins::penguins |> 
  summarise(
    mean_bill_length = mean(bill_length_mm, na.rm = TRUE),
    mean_bill_depth = mean(bill_depth_mm, na.rm = TRUE),
    mean_flipper_length = mean(flipper_length_mm, na.rm = TRUE),
    mean_body_mass = mean(body_mass_g, na.rm = TRUE)
  )
```

Good summarise

```{r}
palmerpenguins::penguins |> 
  summarise(
    across(
      .cols = c(
        bill_length_mm, bill_depth_mm, 
        flipper_length_mm, body_mass_g
      ),
      .fns = \(x) mean(x, na.rm = TRUE)
    )
  )
```

Using multiple functions

```{r}
palmerpenguins::penguins |> 
  summarise(
    across(
      .cols = c(
        bill_length_mm, bill_depth_mm, 
        flipper_length_mm, body_mass_g
      ),
      .fns = c(
        \(x) mean(x, na.rm = TRUE),
        \(x) var(x, na.rm = TRUE)
      )
    )
  )
```

Better column names by naming function in vector

```{r}
palmerpenguins::penguins |> 
  summarise(
    across(
      .cols = c(
        bill_length_mm, bill_depth_mm, 
        flipper_length_mm, body_mass_g
      ),
      .fns = c(
        mean = \(x) mean(x, na.rm = TRUE),
        variance =\(x) var(x, na.rm = TRUE)
      )
    )
  )
```

Better separator for pivot_longer()

```{r}
palmerpenguins::penguins |> 
  summarise(
    across(
      .cols = c(
        bill_length_mm, bill_depth_mm, 
        flipper_length_mm, body_mass_g
      ),
      .fns = c(
        mean = \(x) mean(x, na.rm = TRUE),
        variance =\(x) var(x, na.rm = TRUE)
      ),
      .names = '{.col}----{.fn}'
    )
  ) |> 
  pivot_longer(
    cols = everything(),
    names_sep = '----',
    names_to = c('variable', 'stat')
  )
```

Better approach for many columns using tidyselect helpers

```{r}
modeldata::ames |> 
  summarise(
    across(
      .cols = where(is.numeric),
      .fns = c(
        mean = \(x) mean(x, na.rm = TRUE),
        variance =\(x) var(x, na.rm = TRUE)
      ),
      .names = '{.col}----{.fn}'
    )
  ) |> 
  pivot_longer(
    cols = everything(),
    names_to = c('variable', 'stat'),
    names_sep = '----'
  )
```

Combine multiple select helpers

```{r}
modeldata::ames |> 
  summarise(
    across(
      .cols = where(is.numeric) & starts_with('Lot'),
      .fns = c(
        mean = \(x) mean(x, na.rm = TRUE),
        variance =\(x) var(x, na.rm = TRUE)
      ),
      .names = '{.col}----{.fn}'
    )
  ) |> 
  pivot_longer(
    cols = everything(),
    names_to = c('variable', 'stat'),
    names_sep = '----'
  )
```

You can use across() with mutate as well.

```{r}
modeldata::ames |> 
  mutate(
    across(
      .cols = matches('(Y|y)ear'),
      .fns = \(x) lubridate::make_date(year = x)
    )
  ) |> 
  select(where(lubridate::is.Date))
```
